Data cleaning and processing script

### 2023-03-21 code/modeling_code/PP_7DMA_model.Rmd
model building

```{r}
rmse <- aug_test %>% rmse(truth = case_7dma, .pred)
rsq <- aug_test %>% rsq(truth = case_7dma, .pred)
PP_7DMA_metrics <- full_join(rmse, rsq)

@@

>>>>

```{r}
rmse <- aug_test %>% rmse(truth = case_7dma, .pred)
rsq <- aug_test %>% rsq(truth = case_7dma, .pred)
huber <- aug_test %>% huber_loss(truth = case_7dma, .pred)
PP_7DMA_metrics <- bind_rows(rmse, rsq, huber)

@@

## Plot 7DMA and prediction temporally

```{r}
aug_test %>% ggplot(aes(x=date)) +
  geom_point(aes(y=case_7dma), color = "blue") +
  geom_line(aes(y=case_7dma), color = "blue") +
  geom_point(aes(y=.pred), color = "green") +
  geom_line(aes(y=.pred), color = "green")

## Use model on testing data

```{r}
aug_test2 <- augment(fit_7dma, test_data)
aug_test2 %>% select(case_7dma, .pred)
```

```{r}
aug_test2 %>% ggplot(aes(x=date)) +
  geom_point(aes(y=case_7dma), color = "blue") +
  geom_line(aes(y=case_7dma), color = "blue") +
 geom_point(aes(y=.pred), color = "green") +
  geom_line(aes(y=.pred), color = "green")
```

## Evaluate model performance on testing set

```{r}
rmse <- aug_test2 %>% rmse(truth = case_7dma, .pred)
rsq <- aug_test2 %>% rsq(truth = case_7dma, .pred)
huber <- aug_test2 %>% huber_loss(truth = case_7dma, .pred)
PP_7DMA_test_metrics <- bind_rows(rmse, rsq, huber)

saveRDS(PP_7DMA_test_metrics, here("data/modeling_data/metrics/PP_7DMA_test_performance.RDS"))

PP_7DMA_test_metrics
```

## Use model on testing dates

```{r}
aug_test3 <- augment(fit_7dma, test_dates)
aug_test3 %>% select(case_7dma, .pred)
```

## Evaluate model performance on testing dates

```{r}
rmse <- aug_test3 %>% rmse(truth = case_7dma, .pred)
rsq <- aug_test3 %>% rsq(truth = case_7dma, .pred)
huber <- aug_test3 %>% huber_loss(truth = case_7dma, .pred)
PP_7DMA_BA5_metrics <- bind_rows(rmse, rsq, huber)

saveRDS(PP_7DMA_BA5_metrics, here("data/modeling_data/metrics/PP_7DMA_BA5_performance.RDS"))

PP_7DMA_BA5_metrics
```

```{r}
aug_test3 %>% ggplot(aes(x=date)) +
  geom_point(aes(y=case_7dma), color = "blue") +
  geom_line(aes(y=case_7dma), color = "blue") +
 geom_point(aes(y=.pred), color = "green") +
  geom_line(aes(y=.pred), color = "green")
```

### 2023-03-21 code/processing_code/processing.qmd
model building

```{r}
#N1 data here
n1 <- n1 %>% mutate(ct = as.numeric(ct), 
              copy_num_uL_rxn = as.numeric(10^((ct-36.046)/-3.5293)), #use standard curve slope here
              copy_num_uL_rxn = replace_na(copy_num_uL_rxn, 0.004)) #replace NAs with limit of detection (0.004 copies per uL of the reaction)

#N2 data here (because they have different standard curve equations)
n2 <- n2 %>% mutate(ct = as.numeric(ct),
              copy_num_uL_rxn = as.numeric(10^((ct-37.731)/-3.2505)),
              copy_num_uL_rxn = replace_na(copy_num_uL_rxn, 0.004))

@@

wbe_year2 <- bind_rows(n1, n2) %>% #combine dataframes with same number of rows
  mutate(
    date=as.Date(sample_date, format = "%d-%b-%y"), #change to date format
    facility=substr(sample_id, 1,2), #first two letters in sample_id is facility ID
    biological_replicate=substr(sample_id, nchar(sample_id), nchar(sample_id)), #last number in sample_id is the replicate number
    ct=as.numeric(ifelse(ct=="Undetermined", NA, ct)) #set NAs
  ) %>% 
  arrange(date, facility, target, biological_replicate) %>% 
  select(date, facility, target, biological_replicate, collection_num, ct, copy_num_uL_rxn) %>%
  mutate(facility = as.factor(facility), 
         facility = recode(facility, NO = "A", MI = "B", CC = "C"), 
         facility = ordered(facility, levels = c("A", "B", "C")), #set facility to factor and then set factor levels
         copy_num_L = copy_num_uL_rxn*20/5*60/280*1000*1000) %>% 
  filter(between(date, as.Date("2021-07-06") , as.Date("2022-12-21")))

@@

```{r}
plant_clean <- plant %>% 
  mutate(date = as.Date(date, format = "%m/%d/%Y"), #change to date format
         influent_flow_L = influent_flow_mg*1e6*231*(0.0254^3)*1000) %>%  #convert flow 
  mutate(facility = as.factor(wrf), facility = recode(facility, NO = "A", MI = "B", CC = "C"), facility = ordered(facility, levels = c("A", "B", "C"))) %>% #set facility to factor and then set factor levels
   select(date, facility, influent_flow_L, influent_tss_mg_l)

@@

```{r}
n1_n2_plant <- left_join(wbe_year2_avg, plant_clean, by = c("date","facility")) %>% 
  mutate(total_copies = avg_copy_L * influent_flow_L)

>>>>

n1 \<- n1 %\>% mutate(ct = as.numeric(ct),

copy_num_uL_rxn = as.numeric(10\^((ct-36.046)/-3.5293)), #use standard curve slope here

copy_num_uL_rxn = replace_na(copy_num_uL_rxn, 0.004), #replace NAs with limit of detection (0.004 copies per uL of the reaction)

run_date = as.Date(run_date))

```{r}
#N1 data here
n1 <- n1 %>% mutate(ct = as.numeric(ct), 
              copy_num_uL_rxn = as.numeric(10^((ct-36.046)/-3.5293)), #use standard curve slope here
              copy_num_uL_rxn = replace_na(copy_num_uL_rxn, 0.004), #replace NAs with limit of detection (0.004 copies per uL of the reaction)
              run_date = as.Date(run_date))

#N2 data here (because they have different standard curve equations)
n2 <- n2 %>% mutate(ct = as.numeric(ct),
              copy_num_uL_rxn = as.numeric(10^((ct-37.731)/-3.2505)),
              copy_num_uL_rxn = replace_na(copy_num_uL_rxn, 0.004),
              run_date = as.Date(run_date))

@@

wbe_year2 <- bind_rows(n1, n2) %>% #combine dataframes with same variables
  mutate(
    date=as.Date(sample_date, format = "%d-%b-%y"), #change to date format
    facility=substr(sample_id, 1,2), #first two letters in sample_id is facility ID
    biological_replicate=substr(sample_id, nchar(sample_id), nchar(sample_id)), #last number in sample_id is the replicate number
    ct=as.numeric(ifelse(ct=="Undetermined", NA, ct)) #set NAs
  ) %>% 
  arrange(date, facility, target, biological_replicate) %>% 
  select(date, facility, target, biological_replicate, collection_num, ct, copy_num_uL_rxn) %>%
  mutate(facility = as.factor(facility), 
         facility = recode(facility, NO = "A", MI = "B", CC = "C"), 
         facility = ordered(facility, levels = c("A", "B", "C")), #set facility to factor and then set factor levels
         copy_num_L = copy_num_uL_rxn*20/5*60/280*1000*1000) %>% 
  filter(between(date, as.Date("2021-07-06") , as.Date("2022-12-21")))

@@

```{r}
plant_clean <- plant %>% 
  mutate(date = as.Date(date), #change to date format
         influent_flow_L = influent_flow_mg*1e6*231*(0.0254^3)*1000) %>%  #convert flow from millions of gallons to liters
  mutate(facility = as.factor(wrf), facility = recode(facility, NO = "A", MI = "B", CC = "C"), facility = ordered(facility, levels = c("A", "B", "C"))) %>% #set facility to factor and then set factor levels
   select(date, facility, influent_flow_L, influent_tss_mg_l)

@@

```{r}
n1_n2_plant <- left_join(wbe_year2_avg, plant_clean, by = c("date", "facility")) %>% 
  mutate(total_copies = avg_copy_L * influent_flow_L)
